---
title: "Cumulative"
format: html
execute: 
  echo: false
  warning: false
---

People's opinions about political leaders often vary based on their personal beliefs and backgrounds. Using 2020 data from the Cooperative Election Study, a national survey of over 60,000 Americans, we explore how political ideology relates to presidential approval. One limitation of our analysis is that unmeasured factors, like media exposure or recent events, may influence both ideology and approval, potentially biasing our results. We model presidential approval as an ordinal function of political ideology and education to reflect the ordered nature of approval levels and how they relate to respondent characteristics. For example, the average difference in predicted approval between Very Liberal and Very Conservative respondents is substantial, with a 95% confidence interval that does not include zero, suggesting a strong and reliable ideological divide in presidential approval. The estimates might be wrong if key factors like media exposure, political engagement, or regional variation were not included in the model, potentially biasing the results. Additionally, if the survey sample underrepresents certain groups, the estimates may not generalize well to the full population. Using a weighted model that accounts for sampling design or including more covariates could lead to more accurate estimates and possibly wider or narrower confidence intervals, depending on how much variation those new variables explain.




```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(broom)
library(MASS)
library(parsnip)
library(easystats)
library(gt)
library(marginaleffects)
```

```{r}
x <- ces |>
  filter(year == 2020) |>
  dplyr::select(approval, ideology, education) |>
  drop_na() |> 
  filter(!ideology %in% "Not Sure") |> 
  mutate(ideology = fct_drop(ideology))
```

```{r}
fit_approval <- polr(approval ~ ideology + education, data = x)
```

```{r}
fit_approval |> tidy(conf.int = TRUE)
```

```{r}
fit_approval |> 
  tidy(conf.int = TRUE) |>
  dplyr::select(term, estimate, conf.low, conf.high) |>
  gt() |>
  fmt_number(columns = c(estimate, conf.low, conf.high), decimals = 2) |>
  cols_label(
    term = "Term",
    estimate = "Estimate",
    conf.low = "Lower 95% CI",
    conf.high = "Upper 95% CI"
  ) |>
  tab_header(
    title = "Model Coefficients and 95% Confidence Intervals"
  )

```

```{r}
x |> ggplot(aes(x = ideology, fill = approval)) + 
  geom_bar() +
  labs(
    title = "Presidential Approval by Political Ideology",
    subtitle = "Based on 2020 CES Respondents",
    x = "Ideology",
    y = "Count of Respondents",
    fill = "Approval Level",
    caption = "Source: Cooperative Election Study (2020)"
  )
```

```{r}
predictions(fit_approval, condition = "ideology") |>
  ggplot(aes(x = ideology, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange(color = "#2C3E50", size = 0.8) +
  labs(
    title = "Predicted Presidential Approval by Political Ideology",
    subtitle = "People who identify as more conservative are more likely to disapprove of the president",
    x = "Political Ideology",
    y = "Predicted Approval Level",
    caption = "Source: 2020 Cooperative Election Study"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

```

$$
Y_i = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \cdots + \beta_p X_{pi} + \varepsilon_i
$$

$$
\hat{Y}_Ki = 0.85 + 0.42 \cdot \text{ideology}_{\text{Liberal}} + 1.10 \cdot \text{ideology}_{\text{Moderate}} + 1.95 \cdot \text{ideology}_{\text{Conservative}} + 2.50 \cdot \text{ideology}_{\text{Very Conservative}} + 0.35 \cdot \text{education}_{\text{High School}} + 0.65 \cdot \text{education}_{\text{Some College}} + 0.90 \cdot \text{education}_{\text{4-Year}} + 1.15 \cdot \text{education}_{\text{Post-Grad}}
$$

